State of the Practice for Lattice Boltzmann Method Software

### 2022-02-02 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Initial edits to Section 5 (Comparison of Artifacts to Research Software in General)

Answers \rqref{RQ_CompareArtifacts}.  The software packages were examined for
the presence of artifacts, which were then categorized by frequency. We have
grouped them into common, less common, and rare artifacts in
Table~\ref{artifactspresent}. Common artifacts were found in 16 to 24 ($>$63\%)
of the software packages. Less common artifacts were found in 8 to 15 (30-63\%)
of the software packages. Rare artifacts were found in 1 to 7 ($<$30\%) of the
software packages.

\begin{table}
\begin{center}
\begin{tabular}{ p{5.3 cm} p{4.9 cm} p{5 cm}}
\toprule
\textbf{Common} & \textbf{Less Common} & \textbf{Rare} \\
\midrule
Authors / Developers List & Change Log / Release Notes & API Documentation\\
Bug Tracker & Design Documentation & Developer / Contributor Manual\\
Dependency Notes & Functional Spec. / Notes & FAQ / Forum\\
Installation Guide / Instructions & Performance Information & Verification and
Validation Plan\\
Requirements Spec. / Theory Notes & Test Plan / Script / Cases & Video Guide
(including YouTube)\\
List of Related Publications & User Manual/Guide & \\
Makefile / Build File & Version Control & \\
README File & & \\
License & & \\
Tutorial & & \\
\bottomrule
\end{tabular}
\caption{Artifacts Present in Surveyed Packages, Classified by Frequency}
\label{artifactspresent}
\end{center}
\end{table}

All of the top four AHP ranked packages, ESPREsSo, Ludwig, Palabos, and OpenLB,
have each of the commonly found artifacts, except only three of them have a
requirements specification or theory notes. \wss{Change to clarify that it is
theory notes, not a requirements specification.  Remove requirements spec from
table and put it as rare.}  Palabos is the only one of these four packages that
does not have an artifact from this category.  \wss{** Requirements **}
\wss{Discuss requirements here. Extra information from Peter: I interpreted it
to mean what the software should do. As noted in my report, this was not common,
especially finding any formal document like an SRS. I do not recall any package
having an SRS. Some packages did note in their manuals or artifacts some of the
points that would be found in an SRS document. For example, DL\_MESO notes
functional (``The following is a list of the features that DLMESO currently
supports....'') and non-functional (should work in both ``serial and parallel
running; it can be run on standalone machines, clusters and supercomputers..'')
requirements in its documentation (more on this below). Yes, there was often
some sort of a theory resource. Often it was simply a link or reference to the
theory used in the package. There was significant discrepancy in how much theory
was documented or referenced. Meanwhile, requirements were rare. A few packages
did note some of this information within other artifacts but it was not
comprehensive, and it was not presented in a SRS document. DL\_MESO has some of
the information that belongs in an SRS in its user manual. It is not
comprehensive or presented as it would be in an SRS, but it was probably the
best example of requirements. It has a nice subsection on Functionality.
\citet{Nguyen-HoanEtAl2010} showed requirements are the least commonly produced
type of documentation for scientific software in general.  LBM matches this
trend.}

@@

@@

The top four AHP ranked packages have most of the less common artifacts. At the
time of data collection, only one of the four packages (Palabos) did not have a
user manual or guide, but there was a broken link on the package website
indicating that such an artifact might exist. This broken link was later fixed,
but this is not reflected in our data because it was not present at the time of
data collection. Despite the broken link, Palabos does have a detailed and
informative website. Another one (Ludwig) of the top four packages does not
appear to have publicly visible design documentation. A third package (OpenLB)
from the list does not appear to use a version control system. It is possible
that such a system is used by the package, since its website notes package
version numbers, but the artifacts do not explicitly state the use of such a
system.

@@

@@

We analyzed the state of the practice of software development in the Lattice
Boltzmann Methods software domain by quantitatively and qualitatively measuring,
and comparing, 24 software packages along quality attributes. The software
qualities that were assessed in this report are listed in Section
~\ref{softwarequalities}. A methodology for assessing the state of the practice
of software development in SCS domains was presented. Domain packages were
assessed to answer the software development related research questions listed in
Section~\ref{secIntro} to understand how software quality is impacted by software
development choices, including principles, processes, and tools. Software
developers were interviewed to identify development pain points, and to identify
how software quality is ensured. Quantitative data was used to rank the software
packages using the AHP. The ranking designations were compared with rankings
from the software development community, and we found that many of our top 10
ranked packages are ranked highly by a domain expert. Recommendations for
improving software along quality metrics were made, and highlights are presented
in Section~\ref{highlightedrecommendations} of this conclusion. The findings of
this report can be used to guide future development of SCS, specifically along
quality attributes, and to reduce software quality failures.

>>>>

In this section we answers \rqref{RQ_CompareArtifacts} by comparing the
artifacts that we observed in LBM repositories to those observed and recommended
for research software in general.  roadmap?

As part of filling in the measurement template the LBM packages were examined
for the presence of artifacts, which were then categorized by frequency. We have
grouped them into common, less common, and rare artifacts in
Table~\ref{artifactspresent}. Common artifacts were found in 16 to 24 ($>$63\%)
of the software packages. Less common artifacts were found in 8 to 15 (30-63\%)
of the software packages. Rare artifacts were found in 1 to 7 ($<$30\%) of the
software packages.

\begin{table}
\begin{center}
\begin{tabular}{ p{5.3 cm} p{4.9 cm} p{5 cm}}
\toprule
\textbf{Common} & \textbf{Less Common} & \textbf{Rare} \\
\midrule
Authors / Developers List & Change Log / Release Notes & API Documentation\\
Issue Tracker & Design Documentation & Developer / Contributor Manual\\
Dependency Notes & Functional Spec. / Notes & FAQ / Forum\\
Installation Guide / Instructions & Performance Information & Verification and
Validation Plan\\
Theory Notes & Test Plan / Script / Cases & Video Guide
(including YouTube)\\
List of Related Publications & User Manual/Guide & Requirements Spec.\\
Makefile / Build File & Version Control & \\
README File & & \\
License & & \\
Tutorial & & \\
\bottomrule
\end{tabular}
\caption{Artifacts Present in LBM Packages, Classified by Frequency}
\label{artifactspresent}
\end{center}
\end{table}

All of the top four AHP ranked packages, ESPREsSo, Ludwig, Palabos, and OpenLB,
have each of the commonly found artifacts, except Palabos does not have any
theory notes.  The top four AHP ranked packages have most of the less common
artifacts. At the time of data collection, only one of the four packages
(Palabos) did not have a user manual or guide, but there was a broken link on
the package website indicating that such an artifact might exist. This broken
link was later fixed, but this is not reflected in our data because it was not
present at the time of initial data collection (mid 2020). Despite the broken
link, Palabos does have a detailed and informative website. Another one (Ludwig)
of the top four packages does not appear to have publicly visible design
documentation. A third package (OpenLB) from the list does not appear to use a
version control system. It is possible that such a system is used by the
package, since its website notes package version numbers, but the artifacts do
not explicitly state the use of such a system.

\wss{Change to clarify that it is theory notes, not a requirements
specification.  Remove requirements spec from table and put it as rare.}

\wss{** Requirements **}
\wss{Discuss requirements here. Extra information from Peter: I interpreted it
to mean what the software should do. As noted in my report, this was not common,
especially finding any formal document like an SRS. I do not recall any package
having an SRS. Some packages did note in their manuals or artifacts some of the
points that would be found in an SRS document. For example, DL\_MESO notes
functional (``The following is a list of the features that DLMESO currently
supports....'') and non-functional (should work in both ``serial and parallel
running; it can be run on standalone machines, clusters and supercomputers..'')
requirements in its documentation (more on this below). Yes, there was often
some sort of a theory resource. Often it was simply a link or reference to the
theory used in the package. There was significant discrepancy in how much theory
was documented or referenced. Meanwhile, requirements were rare. A few packages
did note some of this information within other artifacts but it was not
comprehensive, and it was not presented in a SRS document. DL\_MESO has some of
the information that belongs in an SRS in its user manual. It is not
comprehensive or presented as it would be in an SRS, but it was probably the
best example of requirements. It has a nice subsection on Functionality.
\citet{Nguyen-HoanEtAl2010} showed requirements are the least commonly produced
type of documentation for scientific software in general.  LBM matches this
trend.}

@@

 \wss{Kelly paper about how researchers document their theory.  Close to being enough - just need traceability and other things.}

@@

@@

\wss{README files discussed in PranaEtAl2018.}

\wss{At some point should get into the intended use of the software.  The
artifacts will change depending on the intention for the software.  Maybe just
say at the start that writing as if all software is intended for research
readiness?}

@@

We analyzed the state of the practice of software development in the Lattice
Boltzmann Methods software domain by quantitatively and qualitatively measuring,
and comparing, 24 software packages along quality attributes. The software
qualities that were assessed in this report are listed in
Section~\ref{softwarequalities}. A methodology for assessing the state of the
practice of software development in SCS domains was presented. Domain packages
were assessed to answer the software development related research questions
listed in Section~\ref{secIntro} to understand how software quality is impacted
by software development choices, including principles, processes, and tools.
Software developers were interviewed to identify development pain points, and to
identify how software quality is ensured. Quantitative data was used to rank the
software packages using the AHP. The ranking designations were compared with
rankings from the software development community, and we found that many of our
top 10 ranked packages are ranked highly by a domain expert. Recommendations for
improving software along quality metrics were made, and highlights are presented
in Section~\ref{highlightedrecommendations} of this conclusion. The findings of
this report can be used to guide future development of SCS, specifically along
quality attributes, and to reduce software quality failures.
